Historical Soil Organic Carbon Budget

### 2020-03-19 ResultNotebooks/results_socbudget.Rmd
rewising bennis comments, including docx, including newest resultNotebooks (with first plot)
update method part

---
title: "First SOC budget analysis"
output: html_notebook
---

@@

setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/", forcecache = TRUE)
setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")

@@

```{r plot reg stocks}

library(mip)

@@

```{r Validation Densities}
    cell_GSOC      <- readSource("GSOC",  convert="onlycorrect")
    cell_WISE      <- readSource("WISE",  convert="onlycorrect")
    cell_SoilGrids <- readSource("SoilGrids",  convert="onlycorrect")

```
```{r plot density maps}

@@

>>>>

---
title: "First SOC budget analysis"
output:
  html_document:
    df_print: paged
  pdf_document: default
---

@@

setConfig(cachefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/cache/rev4.18.cellpanda/", forcecache = TRUE)
#setConfig(sourcefolder = "/home/kristine/mnt/rd3mod/inputdata/sources/")
setConfig(sourcefolder = "/home/kristine/WORK/R libraries/rd3mod_inputdata/sources/")

@@

```{r plot reg stock, out.width="100%", out.height=1000}

library(mip)
library(quitte)
library(ggplot2)

@@

```{r table dens climatezones}
    cell_GSOC      <- readSource("GSOC",  convert="onlycorrect")
    cell_WISE      <- readSource("WISE",  convert="onlycorrect")
    cell_SoilGrids <- readSource("SoilGrids",  convert="onlycorrect")

    koeppen     <- readSource("Koeppen", subtype="cellular", convert="onlycorrect" )
    getNames(koeppen) <- tolower(getNames(koeppen))
    koepp2IPCC  <- toolGetMapping("mapping_koeppen_ipcc.csv", type = "sectoral")
    koepp2IPCC$koeppen_geiger <- tolower(koepp2IPCC$koeppen_geiger)
    climateIPCC <- toolAggregate(koeppen, rel=koepp2IPCC, from="koeppen_geiger", to="ipcc_reduced", dim=3)
    climateIPCC <- climateIPCC[,getYears(SoilCarbon),]

    SOCDens_lut <- dimSums(SoilCarbon, dim=3.2) 
    SoilCshare  <- SOCDens_lut[,,"crop"]/collapseNames(SOCDens_lut[,,"natveg"])
    SoilCshare[is.na(SoilCshare)]       <- 1
    SoilCshare[is.infinite(SoilCshare)] <- 1

    SoilCshare_ipcczone <- dimSums(climateIPCC * SoilCshare * Landuse[,,"crop"], dim=1)/dimSums(climateIPCC * Landuse[,,"crop"], dim=1)
    print(SoilCshare_ipcczone)

@@

```{r plot density maps}

cropland_share <- dimSums(SoilCarbon[,,"crop"], dim=3)/dimSums(SoilCarbon[,,"natveg"], dim=3)

luplot::plotmap(cropland_share[,"y2000",])

### 2020-03-19 SOCbudgetPaper/SOCbudget.Rmd
update method part

  Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
output:
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
  rticles::copernicus_article: default
  md_document:
    variant: markdown_github
---

```{r child = 'introduction.Rmd'}
```

```{r child = 'methods.Rmd'}
```

```{r child = 'results.Rmd'}
```

```{r child = 'discussion.Rmd'}
```

```{r child = 'conclusion.Rmd'}
```

>>>>

  Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
output:
  rticles::copernicus_article: default
  bookdown::word_document2: default
  md_document:
    variant: markdown_github
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
    md_extensions: +raw_attribute
---
\newpage

```{r child = 'introduction.Rmd'}
```
\newpage

```{r child = 'methods.Rmd'}
```
\newpage

```{r child = 'results.Rmd'}
```
\newpage

```{r child = 'discussion.Rmd'}
```
\newpage

```{r child = 'conclusion.Rmd'}
```
\newpage

### 2020-03-19 SOCbudgetPaper/SOCbudget.tex
update method part

\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstenst@pik-potsdam.de}{\nolinkurl{kristine.karstenst@pik-potsdam.de}})}

@@

\hypertarget{method-50}{%
\section{Method (50)}\label{method-50}}

@@

\hypertarget{steady-state-soc-stocks-and-decay-rates}{%
\subsubsection{Steady-state SOC stocks and decay rates}\label{steady-state-soc-stocks-and-decay-rates}}

In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{eq}\) are given by
\begin{equation}
SOC^{eq} =\frac{C^{\textrm{in}}}{k}
\label{eq:inoutflow}
\end{equation}
with \(C^{\textrm{in}}\) being carbon inputs to the soil and \(k\) denoting the soil organic carbon decay rate. We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which assume three soil carbon sub-pools (active, slow and passive) and entangled dynamics between them. Carbon inflow to each sub-pool (see \ref{sec:carboninputs}) and decay rates (see \ref{sec:tier2}) of each sub-pool are still the key components to determining steady-state SOC stocks.

@@

We account for different carbon inputs sources depending on the land use type (see table \ref{tab:datasourceinputs}). Following the IPCC methodology carbon inputs are disaggregated into different structural components depending on their lignin and nitrogen content (see @ref(ipcc\_2019\_2019)). For each structural components the sum over all carbon inputs sources is allocated to the respective SOC sub-pools. This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow. Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \ref{tab:datasourceinputs}.

@@

\begin{equation}
k_{sub,t,i} = \quad k_{sub} \quad \cdot temp_{t,i} \quad \cdot wat_{t,i} \quad \underbrace{\cdot till_{t,i} \quad \overbrace{\cdot sf_{t,i}.}^\text{only for sub = active}}_{\text{for sub = (active,slow)}}
\label{eq:decayrates}
\end{equation}

For cropland we performed an assessment of tillage types and irrigation conditions, whereas on pastures and natural vegetation, we assume rainfed and non-tilled conditions. Data sources as well as considered effects for each land use types are shown in table \ref{tab:datasourcedecay}. To account for variations of decay rates within grid cells based on different tillage and irrigation regimes, average rates based on area shares are calculated.

@@

\hypertarget{soc-transfer-between-land-use-types}{%
\subsubsection{SOC transfer between land use types}\label{soc-transfer-between-land-use-types}}

We calculate SOC stocks based on the area shares of land use types (lut) within the half-degree grid cells (i). If land is converted from one land use type into another, a respective share of the SOC stocks have to be reallocated as well. We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via

@@

with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area expansion for a given land use type \(lut\). Note that \(!lut\) denotes the sum over all other land use types, which decreases in the specific time step \(t\). Data sources and methodology on land use states and changes are descripted in \ref{sec:landuse}.

\hypertarget{total-soc-stocks}{%
\subsubsection{Total SOC stocks}\label{total-soc-stocks}}

Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the calculate steady-state stock \(SOC^{eq}\) for each land-use types (lut), each sub-pool (sub) and each annual time step (t) as represented in equation \eqref{eq:steadystate}.

\begin{equation}
SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot \Delta t \qquad \forall lut, sub, i.
\label{eq:steadystate}
\end{equation}

@@

\begin{equation}
SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{land use type specific SOC stock within a grid cell}}}_{\text{total SOC stock within a grid cell}}
\label{eq:totalstock}
\end{equation}

@@

@@

\hypertarget{results}{%
\section{Results}\label{results}}

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

@@

\begin{itemize}
\item
  Carbon displacement via leaching and erosion is neglected in this study.
\item
  Non-net/Gross land use transitions are not tracked in this study.
  \newpage
\end{itemize}

>>>>

\correspondence{Kristine\ Karstens\ (kristine.karstenst@pik-potsdam.de)}

@@

\section{Method (50)}

@@

\subsubsection{Steady-state SOC stocks and decay rates}

In a simple first order kinetic approach the steady-state soil organic
carbon stocks \(SOC^{eq}\) are given by \begin{equation}
SOC^{eq} =\frac{C^{\textrm{in}}}{k}
\label{eq:inoutflow}
\end{equation} with \(C^{\textrm{in}}\) being carbon inputs to the soil
and \(k\) denoting the soil organic carbon decay rate. We use for our
calculations the steady-state method of the refinement of the IPCC
guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which
assume three soil carbon sub-pools (active, slow and passive) and
entangled dynamics between them. Carbon inflow to each sub-pool (see
@ref(sec:carboninputs)) and decay rates (see @ref(sec:tier2)) of each
sub-pool are still the key components to determining steady-state SOC
stocks.

@@

We account for different carbon inputs sources depending on the land use
type (see table @ref(tab:datasourceinputs)). Following the IPCC
methodology carbon inputs are disaggregated into different structural
components depending on their lignin and nitrogen content (see
@ref(ipcc\_2019\_2019)). For each structural components the sum over all
carbon inputs sources is allocated to the respective SOC sub-pools. This
implies that not only the amount of carbon, but also their structural
composition is determining the effective inflow. Data sources for all
considered carbon inputs as well as for lignin and nitrogen content can
be found in table @ref(tab:datasourceinputs).

@@

\begin{equation}
\begin{aligned}
& k_{active,t,i}  & = &~ k_{active}  ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ & \cdot~ sf_{t,i}\\
& k_{slow,t,i}    & = &~ k_{slow}    ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ &\\
& k_{passive,t,i} & = &~ k_{passive} ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ & ~ &
\label{eq:decayrates}
\end{aligned}
\end{equation}

For cropland we performed an assessment of tillage types and irrigation
conditions, whereas on pastures and natural vegetation, we assume
rainfed and non-tilled conditions. Data sources as well as considered
effects for each land use types are shown in table
@ref(tab:datasourcedecay). To account for variations of decay rates
within grid cells based on different tillage and irrigation regimes,
average rates based on area shares are calculated.

@@

\subsubsection{SOC transfer between land use types}

We calculate SOC stocks based on the area shares of land use types (lut)
within the half-degree grid cells (i). If land is converted from one
land use type into others (!lut), a respective share of the SOC stocks
have to be reallocated as well. We account for land conversion at the
beginning of each time step \(t\) by calculating a preliminary stock
\(SOC_{lut,t*}\) via

@@

with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area
expansion for a given land use type \(lut\). Note that \(!lut\) denotes
the sum over all other land use types, which decreases in the specific
time step \(t\). Data sources and methodology on land use states and
changes are described in @ref(sec:landuse).

\subsubsection{Total SOC stocks}

Carbon stocks \(SOC\) for each sub-pool (sub) converge towards the
calculate steady-state stock \(SOC^{eq}\) for each land-use types (lut),
each sub-pool (sub) and each annual time step (t) as represented in
equation @ref(eq:steadystate).

\begin{equation}
SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall lut, sub, i.
\label{eq:steadystate}
\end{equation}

@@

\begin{equation}
SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{$SOC_{lut, t, i}$ - land use type specific SOC stock within cell}}}_{\text{$SOC_{t, i}$ - total SOC stock within cell}}
\label{eq:totalstock}
\end{equation}

@@

\hypertarget{sec:tier1}{%
\subsection{Carbon Budget following Tier 1 (150)}\label{sec:tier1}}

Following the tier 1 approach of the IPCC guidelines vol.~4
(\citet{ipcc_2006_2006}), stocks are estimated via stock change factors
given by the IPCC for the topsoil (0-30 cm) and based on measurement
data and expert choice. The factors differentiate between different crop
and management systems reflecting different dynamics under changed in-
and outflows without explicitly tracking these. This method is used
additionally to the Tier-2 approach to evaluate model accuracy.

\hypertarget{sec:agrimanagement}{%
\subsection{Agricultural management (50)}\label{sec:agrimanagement}}

\hypertarget{sec:landuse}{%
\subsubsection{Landuse and Landuse Change (150)}\label{sec:landuse}}

Our land use patterns are based on the Land-Use Harmonization 2 (LUH2,
\citep{LUH2}) data set, which we aggregate from quarter degree to half
degree resolution. To transform the five different cropland
subcategories of LUH2 into our crop functional types we define country
and year specific area shares based on FAOSTAT data (\citep{FAOSTAT})
(see @ref(appendix:fao2luh2mag). Where no data is available we fill data
gaps by using similar countries as proxies or global averages. Land use
transitions are calculated as area differences of the land use data,
allowing any land use type to either expand or shrink in each time step.

\subsubsection{Crop Production and Residues (300)}

Using half-degree yield data from LPJmL (\citep{LPJmL4_1}) as well as
half-degree cropland pattern (see @ref(\#sec:landuse)) we compile crop
functional type specific half-degree production patterns. We calibrate
them to country level FAOSTAT production (\citep{FAOSTAT}) values to be
in line with historical observations.

Starting from this harvested fraction we calculate above ground and
below ground residual biomass using harvest indices and shoot:root
ratios as described in (\citep{bodirsky2012}). We assume that all below
ground biomass is left on the fields, whereas above ground biosmass can
be drawn away for other purposes such as feeding animals
(\citep{weindl}), fuel or for material uses. We calculate recycle shares
by subtracting all residue usages (see \citep{bodirsky1012}) from the
total above ground biomass.

To transform dry matter values into carbon quantities used as soil
input, we compiled crop functional type specific carbon to dry matter
(c:dm) ratios for harvested organs and applied a universal c:dm ratio of
0.44 for all above and below ground residue biomass.

\subsubsection{Pasture production}

Using livestock production statistics as well as feed mix assumptions as
describted in (\citep{weindl}) we estimating country specific pasture
production.

Using the same method as for crop production we disaggregate and
calibrate half-degree pasture production pattern.

\subsubsection{Livestock Distribution and Manure Excretion (300)}

Since their is no native, biophysical indicator to disaggregate country
level FAOSTAT livestock production values to half-degree pattern, we use
the following rule based assumptions:

\begin{itemize}
\item
  Based on the development state of a country, we devide poultry, egg
  and monogastric meat production into intensive and extensive systems.
\item
  Extensive production is allocated to urban area share within a country
  based on the assumption that these animals are held in households,
  subsistence or small-holder farming systems that are labour rich.
\item
  Intensive production is distributed within a country using the crop
  production (excluding 2nd generation bioenergy crops) share, assuming
  that feed availability is the most driving factor for livestock
  location.
\item
  We split dairy and ruminated meat production into pasture fed and
  non-pasture fed fractions based on the aggregated feed mix within a
  country
\end{itemize}

\begin{itemize}
\item
  Pasture fed animals are apportioned to half-degree cells according to
  the pasture production shares of that cell within a country.
\item
  Non-pasture fed animals are distributed again using crop production
  (excluding 2nd generation bioenergy crops) shares.
\end{itemize}

We calculate excretions using the nitrogen balance of the livestock
system (\citep{bodirsky1012}) by assuming that every feed intake not
incoorpared in the slaugher mass, has to excreted. Using C:N ratios
(given by \citep[(][]{ipcc_2019_2019}) we estimate carbon content

\subsubsection{Irrigation (100)}

Simple growing period calculations together with irrigation shares of
LUH2v2 are used (BB: you dont like d's, on't you?) to estimate water
effects on decay rates.

\subsubsection{Tillage (100)}

Tillage data sets of {[}Vera, others{]} together with rules are used to
drive tillage effect on decay rates. \newpage

\section{Results}

@@

\section{Discussion}

@@

\begin{itemize}
\item
  Carbon displacement via leaching and erosion is neglected in this
  study.
\item
  Non-net/Gross land use transitions are not tracked in this study.
\item
  Within cropland we do not track area transitions, but rather look at
  statistical distributions of the crop functional types. Due to crop
  rotations and missing data on crop specific distributions, these
  transitions would be any way rather uncertain. \newpage
\end{itemize}

### 2020-03-19 SOCbudgetPaper/methods.Rmd
update method part

### Carbon Inputs to the Soil {#sec:carboninputs}
<!-- BB: maybe consider to include as text instead of table. The table takes it a bit out of focus. Maybe you want to have the focus on this rather than on other parts. -->

@@

\begin{equation}
k_{sub,t,i} = \quad k_{sub} \quad \cdot temp_{t,i} \quad \cdot wat_{t,i} \quad \underbrace{\cdot till_{t,i} \quad \overbrace{\cdot sf_{t,i}.}^\text{only for sub = active}}_{\text{for sub = (active,slow)}}
\label{eq:decayrates}
\end{equation}

@@

We calculate SOC stocks based on the area shares of land use types (lut) within the half-degree grid cells (i). If land is converted from one land use type into another, a respective share of the SOC stocks have to be reallocated as well. We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{lut,t*}$ via

<!--
If land is converted from one land use type (lut) into another(!lut)
(BB: like this? then you need to change the order of the !lut in the equation; lut is reduced, !lut is expanded)
KK: Not sure, since it might be three different land use types there is no one into 'the' other.
-->

@@

with $A$ being the area, $AR$ the area reduction and $AE$ the area expansion for a given land use type $lut$. Note that $!lut$ denotes the sum over all other land use types, which decreases in the specific time step $t$. Data sources and methodology on land use states and changes are descripted in \@ref(sec:landuse).

@@

\begin{equation}
SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot \Delta t \qquad \forall lut, sub, i.
\label{eq:steadystate}
\end{equation}

<!-- 
BB: What is delta t doing in this equation? The value woudl be 1, no? 
KK: Yes, but the unit is important, since k is 1/a, we need something with a as unit to counteract it.
We could also just write \cdot 1\unit{a}. I had the feeling the Detla is more elegant.
-->

@@

<!--
## Carbon Budget following Tier 1 (150) {#sec:tier1}

Following the tier 1 approach of the IPCC guidelines vol. 4 (@ipcc_2006_2006), stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on measurements. The factors  differentiate between different crop and management systems reflecting different dynamics under changed in- and outflows without explicitly tracking these. This method is used additionally to the Tier-2 approach to evaluate model accuracy.

## Agricultural management	(50) {#sec:agrimanagement}
We combine data sets to estimate agricultural flows and management decisions on cropland. BB: wow. definetly shorter than 50 words! but do you need text hear at all?

@@

We use the Land-Use Harmonization 2 (LUH2, [@LUH2]) half-degree data for major land-use states. Cropland is subdivided using country scale FAO data ([@FAOSTAT]) as well as crop type (BB: crop functional type? ots not crop-specific) specific half-degree data from LUH2v2 to disaggregate into our 19 crop types on half degree resolution (see appendix for table on crop types and their mapping to FAOSTAT and LUH types as well as for technical discription of the disaggregation method).
Landuse transitions are calculated as area differences, allowing any land use type to either expand or shrink in each time step. Within cropland we do not track area transitions, following the averaging approach as descripted in \@ref(sec:tier2)/\@ref(sec:carbonbudget)
-->
<!-- How to explain why we do not track inside-cropland transitions? --> <!-- BB: just dont explain it. -->

<!--
### Crop Production and Residues	(300)
FAOSTAT production ([@FAOSTAT]) values are combined with feed basket estimates including residue and pasture intake ([@weindl]) and rule based demand shares(BB:?). LPJmL yields and calculated land use patterns (see \@ref(#sec:landuse)) are used to scale down to half-degree.

### Livestock Distribution and Manure Excretion	(300)
Based on [Gridded Livestock of the world] we use rule based assumptions(BB?) to estimate livestock and manure distribution on global scale(BB: more detail). Animal waste system shares are used as in [Bodirsky].

@@

### Tillage (100)
Tillage data sets of [Vera, others] together with rules are used to drive tillage effect on decay rates.
-->

>>>>

### Carbon Inputs to the Soil {#sec:carboninputs}

@@

\begin{equation}
\begin{aligned}
& k_{active,t,i}  & = &~ k_{active}  ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ & \cdot~ sf_{t,i}\\
& k_{slow,t,i}    & = &~ k_{slow}    ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ &\cdot~ till_{t,i} ~ &\\
& k_{passive,t,i} & = &~ k_{passive} ~ &\cdot~ temp_{t,i} ~ &\cdot~ wat_{t,i} ~ & ~ &
\label{eq:decayrates}
\end{aligned}
\end{equation}

@@

We calculate SOC stocks based on the area shares of land use types (lut) within the half-degree grid cells (i). If land is converted from one land use type into others (!lut), a respective share of the SOC stocks have to be reallocated as well. We account for land conversion at the beginning of each time step $t$ by calculating a preliminary stock $SOC_{lut,t*}$ via

@@

with $A$ being the area, $AR$ the area reduction and $AE$ the area expansion for a given land use type $lut$. Note that $!lut$ denotes the sum over all other land use types, which decreases in the specific time step $t$. Data sources and methodology on land use states and changes are described in \@ref(sec:landuse).

@@

\begin{equation}
SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot 1\unit{a} \qquad \forall lut, sub, i.
\label{eq:steadystate}
\end{equation}

@@

## Carbon Budget following Tier 1 (150) {#sec:tier1}

Following the tier 1 approach of the IPCC guidelines vol. 4 (@ipcc_2006_2006), stocks are estimated via stock change factors given by the IPCC for the topsoil (0-30 cm) and based on measurement data and expert choice. The factors  differentiate between different crop and management systems reflecting different dynamics under changed in- and outflows without explicitly tracking these. This method is used additionally to the Tier-2 approach to evaluate model accuracy.

## Agricultural management	(50) {#sec:agrimanagement}

@@

Our land use patterns are based on the Land-Use Harmonization 2 (LUH2, [@LUH2]) data set, which we aggregate from quarter degree to half degree resolution. To transform the five different cropland subcategories of LUH2 into our crop functional types we define country and year specific area shares based on FAOSTAT data ([@FAOSTAT]) (see \@ref(appendix:fao2luh2mag). Where no data is available we fill data gaps by using similar countries as proxies or global averages. Land use transitions are calculated as area differences of the land use data, allowing any land use type to either expand or shrink in each time step.

### Crop Production and Residues	(300)
Using half-degree yield data from LPJmL ([@LPJmL4_1]) as well as half-degree cropland pattern (see \@ref(#sec:landuse)) we compile crop functional type specific half-degree production patterns. We calibrate them to country level FAOSTAT production ([@FAOSTAT]) values to be in line with historical observations.

Starting from this harvested fraction we calculate above ground and below ground residual biomass using harvest indices and shoot:root ratios as described in ([@bodirsky2012]). We assume that all below ground biomass is left on the fields, whereas above ground biosmass can be drawn away for other purposes such as feeding animals ([@weindl]), fuel or for material uses. We calculate recycle shares by subtracting all residue usages (see [@bodirsky1012]) from the total above ground biomass.

<!-- Add formular or drawings here? -->

<!-- Ask benni if it is ok to just cide his paper here and not go into detail, how we distribute the residues -->

To transform dry matter values into carbon quantities used as soil input, we compiled crop functional type specific carbon to dry matter (c:dm) ratios for harvested organs and applied a universal c:dm ratio of 0.44 for all above and below ground residue biomass.

<!-- Do we have to explain our choice for the crop functional type specific data. Anyway we have to discuss here a bit, because the overall balance is not clear to me. -->

### Pasture production
Using livestock production statistics as well as feed mix assumptions as describted in ([@weindl]) we estimating country specific pasture production. 
<!-- Is this true? Ask Isabelle again. -->

Using the same method as for crop production we disaggregate and calibrate half-degree pasture production pattern.
<!--As for cropland production we used pasture yields from LPJmL and pasture and rangeland patterns to derive half-degree pasture production patterns. Again country specific pasture production is used to calibrated the pattern.-->

### Livestock Distribution and Manure Excretion	(300)
Since their is no native, biophysical indicator to disaggregate country level FAOSTAT livestock production values to half-degree pattern, we use the following rule based assumptions:

* Based on the development state of a country, we devide poultry, egg and monogastric meat production into intensive and extensive systems.

* Extensive production is allocated to urban area share within a country based on the assumption that these animals are held in households, subsistence or small-holder farming systems that are labour rich.

* Intensive production is distributed within a country using the crop production (excluding 2nd generation bioenergy crops) share, assuming that feed availability is the most driving factor for livestock location.

* We split dairy and ruminated meat production into pasture fed and non-pasture fed fractions based on the aggregated feed mix within a country

<!-- how to spell out that crop is everything else then pasture? -->

* Pasture fed animals are apportioned to half-degree cells according to the pasture production shares of that cell within a country.

* Non-pasture fed animals are distributed again using crop production (excluding 2nd generation bioenergy crops) shares.

<!-- Add formular, drawings or visualisation here? -->

<!-- TILL HERE -->

We calculate excretions using the nitrogen balance of the livestock system ([@bodirsky1012]) by assuming that every feed intake not incoorpared in the slaugher mass, has to excreted. Using C:N ratios (given by [(@ipcc_2019_2019)]) we estimate carbon content

@@

### Tillage (100)
Tillage data sets of [Vera, others] together with rules are used to drive tillage effect on decay rates.
